---
title: "Drought Analysis"
author: "Chris Simoes"
date: "5/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Script setup  
Install packages, start the README.md, set the random seed  
```{r Script setup, echo=FALSE, include=FALSE}
ip <- installed.packages()
packages_used <- c("sommer", "lme4", "dplyr", "workflowr", "here")

ip <- installed.packages()
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}#END packages_used

here::i_am("analysis/droughtAnalysis.Rmd")

source(here::here("code/addToREADME.R"))
addToREADME(paste0("## ", rmarkdown::metadata$title), append=F)
addToREADME(c(date(), ""))

packages_info <- ip[packages_used, c("Package", "Version", "Built")]
addToREADME(c("The packages used in this script are:", "Package, Version, Built"))
apply(packages_info, 1, function(vec) addToREADME(paste(vec, collapse=" ")))
addToREADME("")

random_seed <- 13123178
set.seed(random_seed)
addToREADME(paste("The random seed is", random_seed))
```

## Defining Main Settings and Simulating data

Reading and simulating data. 2 Locations, 12 varieties, 3 rep, 2 treatments, 3 times (days).



```{r main settings, echo = FALSE}
library(dplyr)
# varieties <- c("A1","A2","A3","A4","A5","A6","A7","A8","A9","A10","A11","A12")
# blocks <- c(1,2,3)
# treatments <- c("wattered","drought")
# locations <- c("Mafra", "Chapeco")
# days <- c(30,60,90)

# source(here::here("code/simulateData.R"))



# dataf <- simulateData(varieties, blocks,treatments,locations,days)
# data1 <- dataf%>%filter(location==UQ("Mafra"))
# hist(data1$pheno)

dataf <- read.csv("data/analise_macro_area_foliar_local_1.csv", sep = ",", header = T)
hist(dataf$avg_area_total)

```

## Modeling with 3 different interactions

### Model 1

Model 1 takes treatments and varieties as fixed effect. The interaction treatments per variety is taken as random effect.

```{r model1, echo = TRUE}
library(tidyr)
library(lme4)
library(emmeans)
library(grid)
library(gridBase)

#creating a column named "centered" for avg_area_total
# dataf <- dataf%>%mutate("pheno" = scale(dataf$avg_area_total, scale=TRUE))

model1 <- lmer(avg_area_total~irrigacao+(1|irrigacao:cultivar), data=dataf)
summary(model1)

#getting the random effect
blups_m1 <- ranef(model1)$irrigacao
blups_m1 <- data.frame(blups_m1)
colnames(blups_m1)<-"blups"

#Formating the data for plotting
blup <- blups_m1%>%mutate("tratVar" = rownames(blups_m1))
df <- data.frame(x = blup$tratVar)
dfinal <- df %>% separate(x, c("treatments","variety"), sep = ":")
blup_1_final <- cbind(blup,dfinal)
blup_1_final

min1 <- round(min(blup_1_final$blups)-1,1)
max1 <- round(max(blup_1_final$blups)+1.5,1)

blup_1_final <- blup_1_final[with(blup_1_final, order(treatments,blups)),]
# View(blup_1_final)

# Taking adjusted means. The ideal is take the adjusted mean per treatment.
# adj1 <- summary(lsmeans(model1, "variety"))

# ## Plot, but suppress the labels
midpts <- barplot(blup_1_final$blups, col=rainbow(20), names.arg="", beside= TRUE)

## Use grid to add the labels
vps <- baseViewports()
pushViewport(vps$inner, vps$figure, vps$plot)

grid.text(blup_1_final$varieties,
    x = unit(midpts, "native"), y=unit(-1, "lines"),
    just="right", rot=50)


my_bar <- barplot(blup_1_final$blups , border=F , names.arg=blup_1_final$variety , 
                  las=2 , 
                  col=c(rgb(0.3,0.1,0.4,0.6) , rgb(0.3,0.5,0.4,0.6) , rgb(0.3,0.9,0.4,0.6) ,  rgb(0.3,0.9,0.4,0.6)) , 
                  ylim=c(min1,max1) , 
                  main="" )

# Add abline
# abline(v=5 , col="grey")
 

```


### Model 2

Now, filtering for the drought confdition.
Model 2 takes treatments, blocks and days as fixed effect. Only variety is taken as random effect.

```{r model2, echo=TRUE}
data2 <- dataf%>%filter(irrigacao==UQ("nao_irrigado"))
model2 <- lmer(avg_area_total~ rep + avaliacao + (1|cultivar), data=data2)
summary(model2)
blups_m2 <- ranef(model2)$cultivar


colnames(blups_m2)<-"blups"

#Taking adjusted means
adj2 = coef(model2)$cultivar
#Formating the data for plotting
blup2 <- blups_m2%>%mutate("variety" = rownames(blups_m2), "adjMeans"= adj2[,1])
blup2 <- blup2[with(blup2, order(blups)),]

## Plot, but suppress the labels
midpts <- barplot(blup2$adjMeans, col=rainbow(20), names.arg="")
## Use grid to add the labels    
vps <- baseViewports()
pushViewport(vps$inner, vps$figure, vps$plot)

grid.text(blup2$variety,
    x = unit(midpts, "native"), y=unit(-1, "lines"),
    just="right", rot=50)




```

### Model 3
Here I'm taking the whole data (all tratments).
Model 3 takes blocks as fixed effect, then makes autoregressive residual (AR1) for the days (It means the correlation residual due to measurements have been made on the same plant). Then variety is a random effect combined with treatments.


```{r model3, echo=TRUE}
library(sommer)
# data2 <- data1%>%filter(treatments==UQ("Drought"))

data3<-data.frame(variety=as.factor(dataf$cultivar),
                     # location=as.factor(lColumn),
                     blocks=as.factor(dataf$rep),
                     treatments=as.factor(dataf$irrigacao),
                     days=as.factor(dataf$avaliacao),
                     daysDouble=as.double(dataf$avaliacao),
                     pheno=as.double(dataf$avg_area_total))

model3 <- mmer(pheno~blocks, random=~vs(days, Gu=AR1(days, rho=0.3)) + vs(treatments,variety), rcov=~units, data=data3, verbose = FALSE)
summary(model3)


blups_m3D <- as.data.frame(model3$U$`nao_irrigado:variety`)
blups_m3W <- as.data.frame(model3$U$`irrigado:variety`)
blups_m3 <- rbind(blups_m3D,blups_m3W)
colnames(blups_m3)<-"blups"

#Taking adjusted means
listMeans <- predict.mmer(model3,classify = c("treatments","variety"))
adj3 <- listMeans$pvals$predicted.value
adj3_names <- listMeans$pvals$variety

#preparing column with treatments
varieties <- unique(dataf$cultivar)
trt_Drought <- c(rep("nao_irrigado",length(varieties)))
trt_Wattered <- c(rep("irrigado",length(varieties)))
trt_names <- append(trt_Drought,trt_Wattered)

#Formating the data for plotting
blup3 <- blups_m3%>%mutate("variety" = adj3_names, "adjMeans"=adj3, "treatments"=trt_names)
blup3 <- blup3[with(blup3, order(treatments, adjMeans)),]


# Plot, but suppress the labels
midpts <- barplot(blup3$adjMeans, col=rainbow(20), names.arg="")

# Use grid to add the labels
vps <- baseViewports()
pushViewport(vps$inner, vps$figure, vps$plot)

grid.text(blup3$variety,
    x = unit(midpts, "native"), y=unit(-1, "lines"),
    just="right", rot=50)


```

## Comparing Models

Here I am making a regression with adjusted means between model2 and model3.

```{r regression, echo=TRUE}

# Ordering the data per variety
blup2 <- blup2[with(blup2, order(variety)),]
blup3 <- blup3[with(blup3, order(treatments, variety)),]

plt_dataf = data.frame(blup2Means=as.double(blup2$adjMeans),blup3Means=as.double(blup3$adjMeans[1:length(varieties)]))
m2m3 <- lm(blup2Means ~ blup3Means, data=plt_dataf)
summary(m2m3)
with(plt_dataf,plot(blup3Means, blup2Means))
# abline(m2m3)
coef_m2m3 <- coefficients(m2m3)

r2 <- var(plt_dataf$blup2Means,plt_dataf$blup3Means)/sqrt(var(plt_dataf$blup2Means)*var(plt_dataf$blup3Means))

cat("The coefficent of regression between models 2 and 3 is", round(r2,4),"\n")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
